---
title: "Analysis of scale-dependent biodiversity changes with mobr"
author: "Felix May"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis of scale-dependent biodiversity changes with mobr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installing `mobr`

The package `mobr` is currently available on GitHub and you can freely download 
the source code there.

[mobr on GitHub](https://github.com/MoBiodiv/mobr)

However, when you just want to use the package it is more easy to install the 
package directly from GitHub using the package `devtools`.

```{r, eval = F}
library(devtools)
install_github'MoBiodiv/mobr')
```

# Data structure required by `mobr

To work with `mobr` you need two matrix-like data tables:

1. Species abundance data in a community matrix
2. Attibures of the sampling plots

The community matrix includes counts (= numbers of individuals) for all species
in all sampling plots. The rows represent plots and the columns species.

The table with the plot attributes can include for instance spatial coordinates
of plots, experimental treatments and/or environmental variables.

Throughout this vignette we use an example of a study on the effects of an
invasive on biodiversity.

**More Infos on the example data sets!!!**

This data set is included in `mobr`and available after

```{r, message=FALSE}
library(mobr)
data(inv_comm) # Community matrix
data(inv_plot_attr) # Plot attributes
```

```{r}
str(inv_comm)
head(inv_plot_attr)
```

The plot attributes include the information if a plot is located in invaded or
uninvaded sites as well as the spatial xy coordinates of a plot.

## Preparing data

In order to analyse the data with `mobr` the two data tables have to be combined
into on single object

```{r}
inv_mob_in <- make_mob_in(inv_comm, inv_plot_attr)
```

# Exploratory data analysis

The package `mobr`offers functions for exploratory data analysis and visualization.
The function `mob_stats` calculates the following biodiversity indices for every plot
and for every groups defined by one grouping variable (e.g. experimental treatment)
defined by the user.

* Number of individuals
* Observed species richness
* Rarefied species richness
* Hurlberts PIE (= Probability of Interspecific Encounter) (Hurlbert 1971)

```{r}
inv_stats <- mob_stats(inv_mob_in, group_var = "group")
```

There are also functions for plotting the biodiversity indices for the sampling
plot level and for the group level

**Although these functions need revision - I know**

```{r, fig.width = 7, fig.height = 5}
plot_samples(inv_stats$samples)
```

** This plot indicates we need to do something about plots without individuals!!!**


```{r, fig.width = 7, fig.height = 7}
plot_groups(inv_stats$groups)
```

In random samples PIE does not depend on the sampling effort and samplin scale.
However, if there is aggregation of conspecific individuals in space than the PIE
from a sampling plot will be lower than the PIE from pooled samples. Therefire a 
comparison of plot-base vs. group-based PIE provides and estimate of aggregation
or clumping respectively.

```{r, fig.width = 7, fig.height = 4}
plot_betaPIE(inv_stats)
```

# Full analysis

The complete analysis using `mobr` aims at disentangling the consequences of three
biodiversity components on observed changes in species richness

1. Species abundance distribution
2. Number of individuals
3. Aggregation (clumping) of conspecific individuals


```{r, include=FALSE}
inv_mob_out = get_delta_stats(inv_mob_in, 'group', ref_group='uninvaded',
                              type='discrete', log_scale=TRUE, nperm=100)
```

```{r, eval=FALSE}
inv_mob_out = get_delta_stats(inv_mob_in, 'group', ref_group='uninvaded',
                              type='discrete', log_scale=TRUE, nperm=100)
```


```{r, fig.width=8, fig.height=8}
plot(inv_mob_out, 'invaded', 'uninvaded')
```

Here will be of course text explaining the figure. But I guess there will be 
changes to the figure, so I will wait until the final figure layout is clear. 
Here are my wishes.

1. In the top left panel do nto truncate the curves at the group with
less individuals. I suggest to show all data, because onyl then it gets really
clear that one group includes more individuals

2. Bottom row, middle penal. Without further explanation it is nto clear how the
difference between two curves with the no. of plots as x-axis (middle row,
left and middle panel) switches a no. of individuals x-axis. I think the x-axis
of the bottom row middle panel should be different from no. of individuals.
Something like "Expected no. of individuals" or "scaled no. of individuals"




# References

1.Gotelli, N.J. & Colwell, R.K. (2001). Quantifying biodiversity: procedures and pitfalls in the measurement and comparison of species richness. Ecology letters, 4, 379–391

2.Hurlbert, S.H. (1971). The Nonconcept of Species Diversity: A Critique and Alternative Parameters. Ecology, 52, 577–586


